---
title: "Vignette Siland"
output: html_vignette
author : "Olivier Martin and Florence Carpentier"
date: "`r Sys.Date()`" 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

The package **siland** aims to estimate the scale and th shape of influence for landscape variables on measurements. For this two methods are available.The first one is based on buffer computing around location of obserevd data. Landscape descritpion is based on sig files with sf objects. The second one is based on Spatial Influence Functions (SIF). Landscape descritpion is based on raster discretisation for the landscape. If a sig files is given for argument in this secodn fucntion, then a resterisation procedure is achieved. The SIF fucntion describes how the influence of a pixel in the landscape in the raster decreases with the distance  from the locations of observed data. 

# Data example
 

```
library(siland)
library(dataSiland)
library(landSiland)
```

```{r,echo=FALSE,fig.width=7,fig.height=7}
plot(st_geometry(landSiland))
```
```{r,echo=TRUE,fig.width=7,fig.height=7,eval=FALSE}
sel=landSiland$L1==1
plot(st_geometry(landSiland[sel,]),add=T,col=2)
```
```{r,echo=TRUE,fig.width=7,fig.height=7,eval=FALSE}
sel=landSiland$L2==1
plot(st_geometry(landSiland[sel,]),add=T,col=3)

```
 
# Approach Bsiland


# Approach Fsiland

The **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

# Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
